---
title: "p8105_hw1_smy2122"
author: "Sarah Younes"
date: "2023-09-20"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 0.1

I ran usethis::create_github_token() in the console to get the token to push to GitHub.

## Problem 1

I installed the moderndive package in the console. The code chunk below shows loading the tidyverse package, the moderndive package, and the early January hourly weather dataset.

```{r loading packages and datasets, results = "hide"}
library(tidyverse)
library(moderndive)
data("early_january_weather")
```

The early January hourly weather dataset shows hourly meteorological data throughout early January 2013 for the 3 airports closest to New York City: LGA, JFK, and EWR. The dataset contains `r ncol(early_january_weather)` columns and `r `nrow(early_january_weather)` rows. The columns are the variables, which include the weather station, year, month, day, and hour of recording, temperature ("temp", measured in degrees Fahrenheit), relative humidity ("humid"), wind direction, wind speed, wind gust, precipitation, and date and hour of recording ("time_hour", stored as a single value in seconds as per the POSIXct format). The mean for the temperature is `r mean(pull(early_january_weather, temp))` degrees Fahrenheit.

The code chunk below shows the creation of the scatter plot of temperature by time/hour. Darker points indicate lower humidity and bluer points indicate higher humidity.

```{r ggplot}
ggplot(data = early_january_weather, aes (x = time_hour, y = temp, color = humid)) + geom_point()
```

```{r save, include=FALSE}
ggsave
```

The scatter plot above shows that temperatures are higher during the day and drop somewhat significantly at night. The coldest day is in the beginning of the month around January 2 while the hottest day is in the middle of the month around January 14. On average, as the days go on, both the average temperature and average humidity increase in a positive linear direction from January 1 to around January 16.

# Problem 2

The code chunk below creates a data frame comprised of a random sample of size 10, a logical vector, a character vector, and a factor vector.

```{r problem 2 data frame}
homework_df =
  tibble(
    samp = rnorm(10),
    logical_vec = samp > 0,
    character_vec = c("banana", "strawberry", "mango", "apple", "grape", "kiwi", "blueberry", "passionfruit", "peach", "dragonfruit"),
    factor_vec = factor(c("overly ripe", "ripe", "ripe", "overly ripe", "ripe", "unripe", "unripe", "unripe", "overly ripe", "ripe"))
    )
```

The code chunk below calculates the mean for the random sample, logical vector, character vector, and factor vector.

```{r}
mean(pull(homework_df, samp))
mean(pull(homework_df, logical_vec))
mean(pull(homework_df, character_vec))
mean(pull(homework_df, factor_vec))
```

The mean could only be calculated for the random sample (mean = 0.129) and logical vector (mean = 0.5) variables in the data frame. (Note that, since the random sample changes each time the code is run, the mean for the random sample and logical vector will change accordingly.) However, the character vector and factor vector variables are not numeric or logical and thus return NA.

The next code chunk (which is hidden from the output) converts logical, character, and factor variables from the data frame into numeric vectors.

```{r, results = "hide", warning = FALSE}
as.numeric(pull(homework_df, logical_vec))
as.numeric(pull(homework_df, character_vec))
as.numeric(pull(homework_df, factor_vec))
```

Similar to taking the mean, the logical vector returns numbers of 0 or 1 and the character vectors return NA, which may explain the results for calculating the mean of the logical and character vectors. However, the factor vector now returns values of 1, 2, or 3 that correspond to the order in which levels were assigned, which doe not explain the results for calculating the mean for the factor vector.